### 2024-11-26 paper/paper.qmd
Edits to the paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yan B. Mezhiborsky
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 paper/paper.qmd
Edit to the Data

@@
 library(tidyverse)
-library(palmerpenguins)
+
 ```
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+# 1986
+NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1986[2, ]))
+NBA1986 <- NBA1986[-c(1, 2), ]
+colnames(NBA1986) <- new_col_names
+NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# 1987
+NBA1987 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1987.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1987[2, ]))
+NBA1987 <- NBA1987[-c(1, 2), ]
+colnames(NBA1987) <- new_col_names
+NBA1987[] <- lapply(NBA1987, type.convert, as.is = TRUE)
+
+# 1988
+NBA1988 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1988.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1988[2, ]))
+NBA1988 <- NBA1988[-c(1, 2), ]
+colnames(NBA1988) <- new_col_names
+NBA1988[] <- lapply(NBA1988, type.convert, as.is = TRUE)
+
+# 1989
+NBA1989 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1989.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1989[2, ]))
+NBA1989 <- NBA1989[-c(1, 2), ]
+colnames(NBA1989) <- new_col_names
+NBA1989[] <- lapply(NBA1989, type.convert, as.is = TRUE)
+
+# 1990
+NBA1990 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1990.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1990[2, ]))
+NBA1990 <- NBA1990[-c(1, 2), ]
+colnames(NBA1990) <- new_col_names
+NBA1990[] <- lapply(NBA1990, type.convert, as.is = TRUE)
+
+# 1991
+NBA1991 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1991.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1991[2, ]))
+NBA1991 <- NBA1991[-c(1, 2), ]
+colnames(NBA1991) <- new_col_names
+NBA1991[] <- lapply(NBA1991, type.convert, as.is = TRUE)
+
+# 1992
+NBA1992 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1992.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1992[2, ]))
+NBA1992 <- NBA1992[-c(1, 2), ]
+colnames(NBA1992) <- new_col_names
+NBA1992[] <- lapply(NBA1992, type.convert, as.is = TRUE)
+
+# 1993
+NBA1993 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1993.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1993[2, ]))
+NBA1993 <- NBA1993[-c(1, 2), ]
+colnames(NBA1993) <- new_col_names
+NBA1993[] <- lapply(NBA1993, type.convert, as.is = TRUE)
+
+# 1994
+NBA1994 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1994.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1994[2, ]))
+NBA1994 <- NBA1994[-c(1, 2), ]
+colnames(NBA1994) <- new_col_names
+NBA1994[] <- lapply(NBA1994, type.convert, as.is = TRUE)
+
+# 1995
+NBA1995 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1995.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1995[2, ]))
+NBA1995 <- NBA1995[-c(1, 2), ]
+colnames(NBA1995) <- new_col_names
+NBA1995[] <- lapply(NBA1995, type.convert, as.is = TRUE)
+
+# 1996
+NBA1996 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1996.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1996[2, ]))
+NBA1996 <- NBA1996[-c(1, 2), ]
+colnames(NBA1996) <- new_col_names
+NBA1996[] <- lapply(NBA1996, type.convert, as.is = TRUE)
+
+# 1997
+NBA1997 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1997.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1997[2, ]))
+NBA1997 <- NBA1997[-c(1, 2), ]
+colnames(NBA1997) <- new_col_names
+NBA1997[] <- lapply(NBA1997, type.convert, as.is = TRUE)
+
+# 1998
+NBA1998 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1998.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1998[2, ]))
+NBA1998 <- NBA1998[-c(1, 2), ]
+colnames(NBA1998) <- new_col_names
+NBA1998[] <- lapply(NBA1998, type.convert, as.is = TRUE)
+
+# 1999
+NBA1999 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1999.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1999[2, ]))
+NBA1999 <- NBA1999[-c(1, 2), ]
+colnames(NBA1999) <- new_col_names
+NBA1999[] <- lapply(NBA1999, type.convert, as.is = TRUE)
+
+# 2000
+NBA2000 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2000.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2000[2, ]))
+NBA2000 <- NBA2000[-c(1, 2), ]
+colnames(NBA2000) <- new_col_names
+NBA2000[] <- lapply(NBA2000, type.convert, as.is = TRUE)
+
+# 2001
+NBA2001 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2001.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2001[2, ]))
+NBA2001 <- NBA2001[-c(1, 2), ]
+colnames(NBA2001) <- new_col_names
+NBA2001[] <- lapply(NBA2001, type.convert, as.is = TRUE)
+
+# 2002
+NBA2002 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2002.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2002[2, ]))
+NBA2002 <- NBA2002[-c(1, 2), ]
+colnames(NBA2002) <- new_col_names
+NBA2002[] <- lapply(NBA2002, type.convert, as.is = TRUE)
+
+# 2003
+NBA2003 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2003.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2003[2, ]))
+NBA2003 <- NBA2003[-c(1, 2), ]
+colnames(NBA2003) <- new_col_names
+NBA2003[] <- lapply(NBA2003, type.convert, as.is = TRUE)
+
+# 2004
+NBA2004 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2004.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2004[2, ]))
+NBA2004 <- NBA2004[-c(1, 2), ]
+colnames(NBA2004) <- new_col_names
+NBA2004[] <- lapply(NBA2004, type.convert, as.is = TRUE)
+
+# 2005
+NBA2005 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2005.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2005[2, ]))
+NBA2005 <- NBA2005[-c(1, 2), ]
+colnames(NBA2005) <- new_col_names
+NBA2005[] <- lapply(NBA2005, type.convert, as.is = TRUE)
+
+# 2006
+NBA2006 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2006.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2006[2, ]))
+NBA2006 <- NBA2006[-c(1, 2), ]
+colnames(NBA2006) <- new_col_names
+NBA2006[] <- lapply(NBA2006, type.convert, as.is = TRUE)
+
+# 2007
+NBA2007 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2007.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2007[2, ]))
+NBA2007 <- NBA2007[-c(1, 2), ]
+colnames(NBA2007) <- new_col_names
+NBA2007[] <- lapply(NBA2007, type.convert, as.is = TRUE)
+
+# 2008
+NBA2008 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2008.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2008[2, ]))
+NBA2008 <- NBA2008[-c(1, 2), ]
+colnames(NBA2008) <- new_col_names
+NBA2008[] <- lapply(NBA2008, type.convert, as.is = TRUE)
+
+# 2009
+NBA2009 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2009.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2009[2, ]))
+NBA2009 <- NBA2009[-c(1, 2), ]
+colnames(NBA2009) <- new_col_names
+NBA2009[] <- lapply(NBA2009, type.convert, as.is = TRUE)
+
+# 2010
+NBA2010 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2010.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2010[2, ]))
+NBA2010 <- NBA2010[-c(1, 2), ]
+colnames(NBA2010) <- new_col_names
+NBA2010[] <- lapply(NBA2010, type.convert, as.is = TRUE)
+
+# 2011
+NBA2011 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2011.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2011[2, ]))
+NBA2011 <- NBA2011[-c(1, 2), ]
+colnames(NBA2011) <- new_col_names
+NBA2011[] <- lapply(NBA2011, type.convert, as.is = TRUE)
+
+# 2012
+NBA2012 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2012.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2012[2, ]))
+NBA2012 <- NBA2012[-c(1, 2), ]
+colnames(NBA2012) <- new_col_names
+NBA2012[] <- lapply(NBA2012, type.convert, as.is = TRUE)
+
+# 2013
+NBA2013 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2013.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2013[2, ]))
+NBA2013 <- NBA2013[-c(1, 2), ]
+colnames(NBA2013) <- new_col_names
+NBA2013[] <- lapply(NBA2013, type.convert, as.is = TRUE)
+
+# 2014
+NBA2014 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2014.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2014[2, ]))
+NBA2014 <- NBA2014[-c(1, 2), ]
+colnames(NBA2014) <- new_col_names
+NBA2014[] <- lapply(NBA2014, type.convert, as.is = TRUE)
+
+# 2015
+NBA2015 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2015.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2015[2, ]))
+NBA2015 <- NBA2015[-c(1, 2), ]
+colnames(NBA2015) <- new_col_names
+NBA2015[] <- lapply(NBA2015, type.convert, as.is = TRUE)
+
+# 2016
+NBA2016 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2016.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2016[2, ]))
+NBA2016 <- NBA2016[-c(1, 2), ]
+colnames(NBA2016) <- new_col_names
+NBA2016[] <- lapply(NBA2016, type.convert, as.is = TRUE)
+
+# 2017
+NBA2017 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2017.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2017[2, ]))
+NBA2017 <- NBA2017[-c(1, 2), ]
+colnames(NBA2017) <- new_col_names
+NBA2017[] <- lapply(NBA2017, type.convert, as.is = TRUE)
+
+# 2018
+NBA2018 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2018.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2018[2, ]))
+NBA2018 <- NBA2018[-c(1, 2), ]
+colnames(NBA2018) <- new_col_names
+NBA2018[] <- lapply(NBA2018, type.convert, as.is = TRUE)
+
+# 2019
+NBA2019 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2019.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2019[2, ]))
+NBA2019 <- NBA2019[-c(1, 2), ]
+colnames(NBA2019) <- new_col_names
+NBA2019[] <- lapply(NBA2019, type.convert, as.is = TRUE)
+
+# 2020
+NBA2020 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2020.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2020[2, ]))
+NBA2020 <- NBA2020[-c(1, 2), ]
+colnames(NBA2020) <- new_col_names
+NBA2020[] <- lapply(NBA2020, type.convert, as.is = TRUE)
+
+# 2021
+NBA2021 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2021.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2021[2, ]))
+NBA2021 <- NBA2021[-c(1, 2), ]
+colnames(NBA2021) <- new_col_names
+NBA2021[] <- lapply(NBA2021, type.convert, as.is = TRUE)
+
+# 2022
+NBA2022 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2022.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2022[2, ]))
+NBA2022 <- NBA2022[-c(1, 2), ]
+colnames(NBA2022) <- new_col_names
+NBA2022[] <- lapply(NBA2022, type.convert, as.is = TRUE)
+
+# 2023
+NBA2023 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2023.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2023[2, ]))
+NBA2023 <- NBA2023[-c(1, 2), ]
+colnames(NBA2023) <- new_col_names
+NBA2023[] <- lapply(NBA2023, type.convert, as.is = TRUE)
+
 ```

### 2024-11-29 paper/paper.qmd
Data Cleaning and Variable Creation

@@
 
+# 2024
+NBA2024 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2024.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA2024[2, ]))
+NBA2024 <- NBA2024[-c(1, 2), ]
+colnames(NBA2024) <- new_col_names
+NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
+```
+
+```{r}
+
+# Years range
+years <- 1986:2024
+
+# Normalize columns and calculate MVP index with WS
+for (year in years) {
+  # Construct dataset name dynamically
+  dataset_name <- paste0("NBA", year)
+
+  # Access dataset using `get()`
+  dataset <- get(dataset_name)
+
+  # Check if required columns exist
+  if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
+    # Calculate column-wise averages, including WS
+    averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)
+
+    # Add average columns to the dataset
+    dataset$avg_PTS <- averages["PTS"]
+    dataset$avg_AST <- averages["AST"]
+    dataset$avg_TRB <- averages["TRB"]
+    dataset$avg_STL <- averages["STL"]
+    dataset$avg_BLK <- averages["BLK"]
+    dataset$avg_WS <- averages["WS"]
+
+    # Normalize by averages
+    dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
+    dataset$norm_AST <- dataset$AST / dataset$avg_AST
+    dataset$norm_TRB <- dataset$TRB / dataset$avg_TRB
+    dataset$norm_STL <- dataset$STL / dataset$avg_STL
+    dataset$norm_BLK <- dataset$BLK / dataset$avg_BLK
+    dataset$norm_WS <- dataset$WS / dataset$avg_WS
+
+    # Calculate MVP index, now including WS
+    dataset$MVP_index <- 0.5 * dataset$norm_PTS +
+                         0.2 * dataset$norm_AST +
+                         0.2 * dataset$norm_TRB +
+                         0.1 * dataset$norm_WS  # Adjust weights as needed
+
+    # Assign the modified dataset back to the environment
+    assign(dataset_name, dataset)
+  } else {
+    stop(paste0("One or more required columns (PTS, AST, TRB, STL, BLK, WS) are missing in dataset for year ", year))
+  }
+}
+
+# Create a table of the player with the highest MVP index for each year
+mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), stringsAsFactors = FALSE)
+
+# Iterate over each year to find the top player
+for (year in years) {
+  # Construct dataset name dynamically
+  dataset_name <- paste0("NBA", year)
+
+  # Access dataset using `get()`
+  dataset <- get(dataset_name)
+
+  # Ensure the MVP_index column exists
+  if ("MVP_index" %in% colnames(dataset)) {
+    # Find the row with the highest MVP_index
+    max_row <- dataset[which.max(dataset$MVP_index), ]
+
+    # Extract relevant details (assuming a "Player" column exists)
+    mvp_entry <- data.frame(Year = year, 
+                            Player = max_row$Player, 
+                            MVP_index = max_row$MVP_index,
+                            stringsAsFactors = FALSE)
+
+    # Append to the mvp_table
+    mvp_table <- rbind(mvp_table, mvp_entry)
+  } else {
+    stop(paste0("MVP_index column is missing in dataset for year ", year))
+  }
+}
+
 ```

### 2024-12-01 paper/paper.qmd
Model and Data Edits

@@
 library(tidyverse)
-
+library(dplyr)
+library(janitor)
+library(lubridate)
+library(modelsummary)
+library(rstanarm)
+library(knitr)
+library(bayesplot)
 ```
@@
 NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
+
 ```
@@
     # Calculate MVP index, now including WS
-    dataset$MVP_index <- 0.5 * dataset$norm_PTS +
-                         0.2 * dataset$norm_AST +
-                         0.2 * dataset$norm_TRB +
-                         0.1 * dataset$norm_WS  # Adjust weights as needed
+    dataset$MVP_index <- 0.8 * dataset$norm_PTS +
+                         0.25 * dataset$norm_AST +
+                         0.25 * dataset$norm_TRB +
+                         0.1 * dataset$norm_BLK +
+                         0.1 * dataset$norm_STL +
+                         0.15 * dataset$norm_WS  # Adjust weights as needed
 
@@
 }
+
+nba_list <- list()
+
+for (year in years) {
+  dataset_name <- paste0("NBA", year)
+  dataset <- get(dataset_name)
+  dataset$Year <- year
+  nba_list[[as.character(year)]] <- dataset
+}
+
+nba_master <- do.call(rbind, nba_list)
+write.csv(nba_master, "../data/nba_master.csv", row.names = FALSE)
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+```{r}
+bay_model <- stan_glm(
+  MVP_index ~ PTS + AST + TRB + BLK + STL + WS,
+  data = nba_master,
+   family = gaussian(),
+  prior = normal(0, 5),
+  prior_intercept = normal(0, 5),
+  chains = 4,
+  iter = 2000,
+  warmup = 1000,
+  cores = 4,
+  control = list(adapt_delta = 0.95),
+  seed = 613
+)
+
+```
 

### 2024-12-01 paper/paper.qmd
Edits to model and data

@@
 library(bayesplot)
+library(StanHeaders)
+library(Matrix)
+library(lme4)
+
 ```
@@
 y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
+\mu_i &= \pi_i + \alpha_i + \rho_i + \beta_i + \xi_i + \omega_i \\
+\mu &\sim \mbox{Normal}(0, 2.5) \\
+\pi &\sim \mbox{Normal}(0, 2.5) \\
 \alpha &\sim \mbox{Normal}(0, 2.5) \\
+\rho &\sim \mbox{Normal}(0, 2.5) \\
 \beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\xi &\sim \mbox{Normal}(0, 2.5) \\
+\omega &\sim \mbox{Normal}(0, 2.5) \\
 \sigma &\sim \mbox{Exponential}(1)
@@
 ```{r}
-bay_model <- stan_glm(
-  MVP_index ~ PTS + AST + TRB + BLK + STL + WS,
-  data = nba_master,
-   family = gaussian(),
-  prior = normal(0, 5),
-  prior_intercept = normal(0, 5),
-  chains = 4,
-  iter = 2000,
-  warmup = 1000,
-  cores = 4,
-  control = list(adapt_delta = 0.95),
-  seed = 613
-)
+# Fit a linear regression model
+glm_model <- glm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)
 
+# View the summary of the model
+summary(lm_model)
+
+plot(lm_model)
 ```

### 2024-12-02 paper/paper.qmd
Edits to the paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Finding the True MVP"
+subtitle: "An NBA Analysis Through Linear Regression"
 author: 
@@
 NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
+```
+
+```{r}
+Real_MVP_Data <- read.csv("../NBAMVP/data/01-raw_data/NBA\\ MVPs/Real_MVP_Data.csv")
 
@@
 for (year in years) {
-  # Construct dataset name dynamically
   dataset_name <- paste0("NBA", year)
-
-  # Access dataset using `get()`
   dataset <- get(dataset_name)
 
-  # Check if required columns exist
   if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
-    # Calculate column-wise averages, including WS
     averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)
 
-    # Add average columns to the dataset
     dataset$avg_PTS <- averages["PTS"]
@@
 
-    # Normalize by averages
     dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
@@
 
-    # Calculate MVP index, now including WS
     dataset$MVP_index <- 0.8 * dataset$norm_PTS +
@@
                          0.1 * dataset$norm_STL +
-                         0.15 * dataset$norm_WS  # Adjust weights as needed
+                         0.15 * dataset$norm_WS
 
-    # Assign the modified dataset back to the environment
     assign(dataset_name, dataset)
@@
 # Create a table of the player with the highest MVP index for each year
-mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), stringsAsFactors = FALSE)
+mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), 
+                        PTS = numeric(), AST = numeric(), TRB = numeric(), 
+                        BLK = numeric(), STL = numeric(), WS = numeric(), 
+                        stringsAsFactors = FALSE)
 
@@
 for (year in years) {
-  # Construct dataset name dynamically
   dataset_name <- paste0("NBA", year)
-
-  # Access dataset using `get()`
   dataset <- get(dataset_name)
 
-  # Ensure the MVP_index column exists
   if ("MVP_index" %in% colnames(dataset)) {
-    # Find the row with the highest MVP_index
     max_row <- dataset[which.max(dataset$MVP_index), ]
 
-    # Extract relevant details (assuming a "Player" column exists)
-    mvp_entry <- data.frame(Year = year, 
+    mvp_entry <- data.frame(
+      Year = year, 
       Player = max_row$Player, 
       MVP_index = max_row$MVP_index,
-                            stringsAsFactors = FALSE)
+      PTS = max_row$PTS,
+      AST = max_row$AST,
+      TRB = max_row$TRB,
+      BLK = max_row$BLK,
+      STL = max_row$STL,
+      WS = max_row$WS,
+      stringsAsFactors = FALSE
+    )
 
-    # Append to the mvp_table
     mvp_table <- rbind(mvp_table, mvp_entry)
@@
 nba_master <- do.call(rbind, nba_list)
-write.csv(nba_master, "../data/nba_master.csv", row.names = FALSE)
+write.csv(nba_master, "../data/02-analysis_data/nba_master.csv", row.names = FALSE)
+
+merged_real_mvp <- merge(
+  Real_MVP_Data, 
+  nba_master[, c("Player", "Year", "MVP_index")],  # Select only relevant columns from nba_master
+  by = c("Player", "Year"), 
+  all.x = TRUE
+)
 
@@
 
-plot(lm_model)
 ```
@@
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+ggplot(mvp_table, aes(x = Year, y = MVP_index)) +
+  geom_point(color = "blue", size = 3, alpha = 0.7) +
+  labs(
+    title = "MVP Index Over the Years",
+    x = "Year",
+    y = "MVP Index"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.text = element_text(size = 12),
+    axis.title = element_text(size = 14)
+  )
 ```

### 2024-12-02 paper/paper.qmd
Created some graphs

@@
 library(lme4)
+library(patchwork)
 
@@
 ```{r}
-Real_MVP_Data <- read.csv("../NBAMVP/data/01-raw_data/NBA\\ MVPs/Real_MVP_Data.csv")
+Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")
 
@@
   Real_MVP_Data, 
-  nba_master[, c("Player", "Year", "MVP_index")],  # Select only relevant columns from nba_master
-  by = c("Player", "Year"), 
+  nba_master[, c("Player", "Age", "MVP_index")],  # Select only relevant columns from nba_master
+  by = c("Player", "Age"), 
   all.x = TRUE
 )
+
+merged_real_mvp <- merged_real_mvp %>%
+  mutate(
+    Year = case_when(
+      # If entry matches "YYYY-YY", extract the first 4 characters as the year
+      grepl("^\\d{4}-\\d{2}$", Season) ~ as.numeric(substr(Season, 1, 4)) + 1,
+
+      # If entry matches "Mmm-YY", extract the last 2 characters, convert to 20YY, and add 1
+      grepl("^[a-zA-Z]{3}-\\d{2}$", Season) ~ as.numeric(paste0("20", substr(Season, 5, 6))) + 1,
+
+      # Default: NA for non-matching formats
+      TRUE ~ NA_real_
+    )
+  )
+
+# Convert Year column to numeric
+merged_real_mvp$Year <- as.numeric(merged_real_mvp$Year)
 
@@
 
-ggplot(mvp_table, aes(x = Year, y = MVP_index)) +
+ggplot(merged_real_mvp, aes(x = Year, y = MVP_index)) +
   geom_point(color = "blue", size = 3, alpha = 0.7) +
   labs(
-    title = "MVP Index Over the Years",
+    title = "Scatter: Year vs. MVP Index (Merged Real MVPs)",
     x = "Year",
@@
   ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
-    axis.text = element_text(size = 12),
-    axis.title = element_text(size = 14)
-  )
+  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
+  theme_minimal()
+
+ggplot(mvp_table, aes(x = Year, y = MVP_index)) +
+  geom_point(color = "red", size = 3, alpha = 0.7) +
+  labs(
+    title = "Scatter: Year vs. MVP Index (MVP Table)",
+    x = "Year",
+    y = "MVP Index"
+  ) +
+  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
+  theme_minimal()
+
 ```

### 2024-12-03 paper/paper.qmd
Added Graphs and Rwender Test

@@
 #| message: false
+#| echo: false
 
@@
 library(patchwork)
+library(kableExtra)
 
@@
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
+#| echo: false
+
 # 1986
@@
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
+#| echo: false
+
 Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")
@@
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
+#| echo: false
 
@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
 Talk way more about it.
@@
 # Fit a linear regression model
-glm_model <- glm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)
+lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)
 
@@
 summary(lm_model)
+
+next_year_mvp_stats <- mvp_table %>%
+  filter(Year == 2024) %>%  # Replace with the target year
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+# Predict MVP_index for the next year's MVP
+predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
+print(paste("Predicted MVP Index (mvp_table):", predicted_index_mvp_table))
+
+next_year_real_mvp_stats <- merged_real_mvp %>%
+  filter(Year == 2024) %>%  # Replace with the target year
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+# Predict MVP_index for the next year's MVP
+predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
+print(paste("Predicted MVP Index (merged_real_mvp):", predicted_index_real_mvp))
 
@@
 We can use maths by including latex between dollar signs, for instance $\theta$.
+
+### Model Results
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
+
+# Filter MVPs from 2014 to 2024 for both datasets
+mvp_table_filtered <- mvp_table %>%
+  filter(Year >= 2014 & Year <= 2024) %>%
+  select(Year, Player, MVP_index)
+
+real_mvp_filtered <- merged_real_mvp %>%
+  filter(Year >= 2014 & Year <= 2024) %>%
+  select(Year, Player, MVP_index)
+
+# Predict MVP index for 2025 and add it to both tables
+next_year_mvp_stats <- mvp_table %>%
+  filter(Year == 2024) %>%  # Use 2024 stats as proxy
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+predicted_index_mvp_table <- NA
+if (nrow(next_year_mvp_stats) > 0) {
+  predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
+}
+
+mvp_table_filtered <- bind_rows(
+  mvp_table_filtered,
+  data.frame(
+    Year = 2025,
+    Player = "Predicted MVP",
+    MVP_index = predicted_index_mvp_table
+  )
+)
+
+next_year_real_mvp_stats <- merged_real_mvp %>%
+  filter(Year == 2024) %>%  # Use 2024 stats as proxy
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+predicted_index_real_mvp <- NA
+if (nrow(next_year_real_mvp_stats) > 0) {
+  predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
+}
+
+real_mvp_filtered <- bind_rows(
+  real_mvp_filtered,
+  data.frame(
+    Year = 2025,
+    Player = "Predicted MVP",
+    MVP_index = predicted_index_real_mvp
+  )
+)
+
+# Rename columns for clarity
+mvp_table_filtered <- mvp_table_filtered %>%
+  rename(
+    `Simulated Data - Player` = Player,
+    `Simulated Data - MVP Index` = MVP_index
+  )
+
+real_mvp_filtered <- real_mvp_filtered %>%
+  rename(
+    `Historical Data - Player` = Player,
+    `Historical Data - MVP Index` = MVP_index
+  )
+
+# Join the two tables by Year
+combined_table <- full_join(
+  mvp_table_filtered,
+  real_mvp_filtered,
+  by = "Year"
+)
+
+# Create a styled table using kable
+combined_table %>%
+  kable(
+    format = "html",
+    caption = "Combined MVP Data (2014-2025)",
+    col.names = c(
+      "Year",
+      "Simulated Data - Player",
+      "Simulated Data - MVP Index",
+      "Historical Data - Player",
+      "Historical Data - MVP Index"
+    )
+  ) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
+  add_header_above(c(" " = 1, "Simulated Data" = 2, "Historical Data" = 2))
+
+```
 
@@
 ```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+# Add a source column to the datasets
+mvp_table <- mvp_table %>%
+  mutate(Source = "Simulated MVPs")
+
+merged_real_mvp <- merged_real_mvp %>%
+  mutate(Source = "Historical MVPs")
+
+# Combine datasets
+combined_data <- bind_rows(
+  mvp_table %>% select(Year, PTS, TRB, AST, Source),
+  merged_real_mvp %>% select(Year, PTS, TRB, AST, Source)
 )
+
+# Plot for PTS
+pts_plot <- ggplot(combined_data, aes(x = Year, y = PTS, group = Source, color = Source)) +
+  geom_point(size = 3, alpha = 0.7) +
+  geom_line(alpha = 0.7) +
+  facet_wrap(~ Source) +
+  labs(
+    title = "Year vs. PTS",
+    x = "Year",
+    y = "PTS"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Plot for TRB
+trb_plot <- ggplot(combined_data, aes(x = Year, y = TRB, group = Source, color = Source)) +
+  geom_point(size = 3, alpha = 0.7) +
+  geom_line(alpha = 0.7) +
+  facet_wrap(~ Source) +
+  labs(
+    title = "Year vs. TRB",
+    x = "Year",
+    y = "TRB"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Plot for AST
+ast_plot <- ggplot(combined_data, aes(x = Year, y = AST, group = Source, color = Source)) +
+  geom_point(size = 3, alpha = 0.7) +
+  geom_line(alpha = 0.7) +
+  facet_wrap(~ Source) +
+  labs(
+    title = "Year vs. AST",
+    x = "Year",
+    y = "AST"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Display plots separately
+print(pts_plot)
+print(trb_plot)
+print(ast_plot)
+
 ```
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-12-03 paper/paper.qmd
Edited data section and added references

@@
 library(tidyverse)
+library(rstanarm)
 library(dplyr)
@@
 library(modelsummary)
-library(rstanarm)
 library(knitr)
-library(bayesplot)
 library(StanHeaders)
-library(Matrix)
-library(lme4)
-library(patchwork)
 library(kableExtra)
@@
 
-# Introduction
+# Introduction {#sec-intro}
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR], as well as the following packages to help clean and manipulate our data, [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], [@kableExtra].
+Using these packages we cleaned our data, added years for the data sets that only had seasons, referring to the basketball season, as well as helped create some nice tables, unlike the traditional R software can create.
 
-Overview text
+Our data was collected from [@bbref], we collected data from every player that received a vote for any All-NBA teams since 1986, this includes a vote cast for the First All-NBA Team, the Second All-NBA Team, and the Third All-NBA Team.
+As we believed that taking the players that received votes for the MVP awards would result in a smaller sample.
+We also created a new statistic call `MVP_Index` that will be defined in @sec-variables.
+In addition, we collected a data set for all NBA MVPs since 1986, so we could compare to the ones we have chosen with the `MVP_Index`.
+
+Following @tellingstories, we use techniques and methods from this text we have analyzed the data using models, graphs, tables and other methods of data analysis to conduct our research to find  players that could have been MVPs and what the MVP for the following years will look like(statistically).
+
+## Variables {#sec-variables}
+
+### All NBA Variables
+
+As we collected data from each year since 1986 individually, we have utilized the following variables:
+
+-   `Team:` Whether the player was in the 1st, 2nd, or 3rd All NBA Teams or OVR if they did not make the top three teams but still received a vote.
+-   `Pos:` The player's position including the generalized positions G, F, C, for Guard, Forward, and Center respectively.
+-   `Player:` The name of the NBA player.
+-   `Age:` The age of the NBA player.
+-   `Tm:` The three letter designation for the NBA team the player played for the year.
+-   `G:` Games played by the player.
+-   `MP:` Minutes played per game.
+-   `PTS:` Points per game.
+-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-   `BLK:` Blocks per game
+-   `STL:` Steals Per game
+-   `WS:` The amount of wins allocated to the individual player
+
+As each year has an individual data set we took PTS, TRB, AST, BLK, STL,a nd WS took the average for that year and normalized each player's statistics by dividing them by the average of that year with this we have created the MVP Index which we have defined as:
+
+\$\$
+
+MVP\_{Index} = 0.8PTS\_{Norm} + 0.25AST\_{Norm} + 0.25TRB\_{Norm} + 0.1STL\_{Norm} + 0.1BLK\_{Norm} + 0.15WS\_{Norm}
+
+\$\$
+
+With this `MVP Index` we have created our own list of who could have been MVP base on statistics for that year, essentially comparing how good the player was to othe rplayers in the basketball season.
+
+### Historical MVP Variable
+
+For our second set of date we have all MVPs since 1986 also sourced from [@bbref], and the variables that have been used are the following:
+
+-   `Player:` The name of the NBA player.
+-   `Year:` The year in which the player was selected as MVP.
+-   `Tm:` The three letter designation for the NBA team the player played for the year.
+-   `G:` Games played by the player.
+-   `MP:` Minutes played per game.
+-   `PTS:` Points per game.
+-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-   `BLK:` Blocks per game
+-   `STL:` Steals Per game
+-   `WS:` The amount of wins allocated to the individual player
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+Data collected in sports and more specifically basketball is rather simple.
+When an NBA game is being played there are people who's entire job is to keep track of a players statistics throughout the game, whenever a player makes a shot, or steals they ball it is being logged by that specific person.
+After the standard 82 games, there could be morte or less under different circumstances such ast COVID-19, injuries, or even trades all regular season statistics have been collected.
+As most of our in game statistics are based on the players 'per game' average, the total number of points blocks, etc., are recorded it is divided by the number of games the player played that season.
 
-## Outcome variables
+## Outcome variable: MVP Index
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Our outcome variable is the same as our estimand in @sec-intro.
+As `MVP Index` is dependent on our other statistics, Points, Assists, Rebounds, Steals, Blocks, and Win Shares.
+As we are studying MVPs in our paper this is our Variable of interest ast it represents how good a player was in comparison to the other best players in the NBA that year, from the sample of players that have received any ALL NBA Team vote.
 
@@
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
+```
 
@@
 ### Model Results
+
 ```{r}
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-12-03 paper/paper.qmd
talked about model

@@
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
 # Model
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is twofold.
+Firstly, we will be using a linear regression model to predict the MVP Index of the mvp in the following year regardless of the player.
+This will tell us how much better we can expect the 2025 MVP to be compared to his fellow NBA players.
+Secondly, We will use another data set we have created by selecting players that had the highest MVP index in their respective year whether they had won MVP or not.
+So, Using both these predictions we will be able to compare whether we believe the player with the highest MVP index will in fact be the next NBA MVP.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the Linear analysis model used to investigate what the MVP Index of the next NBA MVP will be. For our model we will use collected data describe in @sec-data.
+
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 ```{r}
-# Fit a linear regression model
+# create a linear regression model
 lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)
@@
 predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
-print(paste("Predicted MVP Index (merged_real_mvp):", predicted_index_real_mvp))
 
@@
 ```{r}
-#| include: true
+#| echo: false
 #| warning: false
 #| message: false
-#| echo: false
+#| label: fig-modeltable
+#| fig-cap: "Historical MVP and Player with highest MVP Index since 2014"
 
@@
     format = "html",
-    caption = "Combined MVP Data (2014-2025)",
     col.names = c(
@@
 #| echo: false
-#| eval: true
 #| warning: false
 #| message: false
+#| label: fig-mvpindex
+#| fig-cap: "MVP index of of Historical and simulated MVPs"
 
-ggplot(merged_real_mvp, aes(x = Year, y = MVP_index)) +
-  geom_point(color = "blue", size = 3, alpha = 0.7) +
+merged_data <- bind_rows(
+  merged_real_mvp %>% mutate(Source = "Historical MVPs"),
+  mvp_table %>% mutate(Source = "Simulated MVPs")
+)
+
+# Create the faceted plot with lines connecting points
+ggplot(merged_data, aes(x = Year, y = MVP_index, group = Source)) +
+  geom_line(aes(color = Source), size = 1, alpha = 0.7) +  # Add lines connecting points
+  geom_point(aes(color = Source), size = 3, alpha = 0.7) +  # Keep the points
+  facet_wrap(~ Source, ncol = 2) +  # Facet by the Source column, side by side
   labs(
-    title = "Scatter: Year vs. MVP Index (Merged Real MVPs)",
+    title = "Scatter: Year vs. MVP Index",
     x = "Year",
@@
   coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
-  theme_minimal()
-
-ggplot(mvp_table, aes(x = Year, y = MVP_index)) +
-  geom_point(color = "red", size = 3, alpha = 0.7) +
-  labs(
-    title = "Scatter: Year vs. MVP Index (MVP Table)",
-    x = "Year",
-    y = "MVP Index"
-  ) +
-  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    legend.position = "none",  # Remove the legend since it's redundant with faceting
+    strip.text = element_text(size = 12, face = "bold")
+  )
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-pts
+#| fig-cap: "MVP PPG over Time"
 
@@
   labs(
-    title = "Year vs. PTS",
     x = "Year",
@@
   labs(
-    title = "Year vs. TRB",
     x = "Year",
@@
   labs(
-    title = "Year vs. AST",
     x = "Year",
@@
 print(pts_plot)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-reb
+#| fig-cap: "MVP RPG over Time"
+
 print(trb_plot)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-ast
+#| fig-cap: "MVP ASG over Time"
+
 print(ast_plot)
-
 ```
@@
 
+```{r}
+# Extract the summary of the model
+lm_summary <- summary(lm_model)
+
+# Extract coefficients as a data frame
+coef_table <- as.data.frame(lm_summary$coefficients)
+colnames(coef_table) <- c("Estimate", "Std. Error", "t value", "Pr(>|t|)")
+
+# Create the kable table
+kable(coef_table, caption = "Linear Regression Model Coefficients") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
+
+print(lm_summary)
+```
+
 ## Posterior predictive check

### 2024-12-03 paper/paper.qmd
Edits to paper

@@
   kable(
-    format = "html",
+    format = "latex",
+    booktabs = TRUE,
     col.names = c(
@@
 ```{r}
-# Extract the summary of the model
 lm_summary <- summary(lm_model)
-
-# Extract coefficients as a data frame
-coef_table <- as.data.frame(lm_summary$coefficients)
-colnames(coef_table) <- c("Estimate", "Std. Error", "t value", "Pr(>|t|)")
-
-# Create the kable table
-kable(coef_table, caption = "Linear Regression Model Coefficients") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
 

### 2024-12-03 paper/paper.qmd
Render terst finally worked !!

@@
 
-\$\$
+$$
 
-MVP\_{Index} = 0.8PTS\_{Norm} + 0.25AST\_{Norm} + 0.25TRB\_{Norm} + 0.1STL\_{Norm} + 0.1BLK\_{Norm} + 0.15WS\_{Norm}
+MVP_{Index} = 0.8PTS_{Norm} + 0.25AST_{Norm} + 0.25TRB_{Norm} + 0.1STL_{Norm} + 0.1BLK_{Norm} + 0.15WS_{Norm}
 
-\$\$
+$$
 
@@
 
--   `Player:` The name of the NBA player.
--   `Year:` The year in which the player was selected as MVP.
--   `Tm:` The three letter designation for the NBA team the player played for the year.
--   `G:` Games played by the player.
--   `MP:` Minutes played per game.
--   `PTS:` Points per game.
--   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
--   `BLK:` Blocks per game
--   `STL:` Steals Per game
--   `WS:` The amount of wins allocated to the individual player
+-`Player:` The name of the NBA player.
+-`Year:` The year in which the player was selected as MVP.
+-`Tm:` The three letter designation for the NBA team the player played for the year.
+-`G:` Games played by the player.
+-`MP:` Minutes played per game.
+-`PTS:` Points per game.
+-`TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-`BLK:` Blocks per game
+-`STL:` Steals Per game
+-`WS:` The amount of wins allocated to the individual player
 
@@
     Year = case_when(
-      # If entry matches "YYYY-YY", extract the first 4 characters as the year
       grepl("^\\d{4}-\\d{2}$", Season) ~ as.numeric(substr(Season, 1, 4)) + 1,
-
-      # If entry matches "Mmm-YY", extract the last 2 characters, convert to 20YY, and add 1
       grepl("^[a-zA-Z]{3}-\\d{2}$", Season) ~ as.numeric(paste0("20", substr(Season, 5, 6))) + 1,
-
-      # Default: NA for non-matching formats
       TRUE ~ NA_real_
@@
 
-```{=tex}
 \begin{align}
@@
 \end{align}
-```
 
 ```{r}
+
 # create a linear regression model
@@
 
-# Create a styled table using kable
-combined_table %>%
-  kable(
-    format = "latex",
-    booktabs = TRUE,
-    col.names = c(
-      "Year",
-      "Simulated Data - Player",
-      "Simulated Data - MVP Index",
-      "Historical Data - Player",
-      "Historical Data - MVP Index"
-    )
-  ) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
-  add_header_above(c(" " = 1, "Simulated Data" = 2, "Historical Data" = 2))
+print(combined_table)
 

### 2024-12-04 paper/paper.qmd
Added results and references

@@
 author: 
-  - Yan B. Mezhiborsky
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Yan Mezhiborsky
+thanks: "Code and data are available at: [https://github.com/Mezhi18/NBAMVP.]"
 date: today
@@
 
-$$
-
-MVP_{Index} = 0.8PTS_{Norm} + 0.25AST_{Norm} + 0.25TRB_{Norm} + 0.1STL_{Norm} + 0.1BLK_{Norm} + 0.15WS_{Norm}
-
-$$
+$$MVP_{Index} = 0.8PTS_{Norm} + 0.25AST_{Norm} + 0.25TRB_{Norm} + 0.1STL_{Norm} + 0.1BLK_{Norm} + 0.15WS_{Norm}$$
 
@@
 
--`Player:` The name of the NBA player.
--`Year:` The year in which the player was selected as MVP.
--`Tm:` The three letter designation for the NBA team the player played for the year.
--`G:` Games played by the player.
--`MP:` Minutes played per game.
--`PTS:` Points per game.
--`TRB:` Total Rebounds per game, including offensive and defensive rebounds.
--`BLK:` Blocks per game
--`STL:` Steals Per game
--`WS:` The amount of wins allocated to the individual player
+-   `Player:` The name of the NBA player.
+-   `Year:` The year in which the player was selected as MVP.
+-   `Tm:` The three letter designation for the NBA team the player played for the year.
+-   `G:` Games played by the player.
+-   `MP:` Minutes played per game.
+-   `PTS:` Points per game.
+-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-   `BLK:` Blocks per game
+-   `STL:` Steals Per game
+-   `WS:` The amount of wins allocated to the individual player
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the MVP Index.
+Then $\pi_i$ as the points per game, then $\alpha_i$ as assists per game, next we have $\rho_i$ as rebounds per game.
+Furthermore, we have $\beta_i$ as blocks per game, then we have $\xi_i$ for steals per game, lastly we have $\omega_i$ for win shares.
+With these variables we will be predicting the MVP Index for the following NBA season.
 
@@
 \mu_i &= \pi_i + \alpha_i + \rho_i + \beta_i + \xi_i + \omega_i \\
-\mu &\sim \mbox{Normal}(0, 2.5) \\
 \pi &\sim \mbox{Normal}(0, 2.5) \\
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+We expect a positive relationship between Between MVP Index and points, assists, rebounds, block, steals and win shares, as as these are all variables that define the MVP Index.
+In addition we expect a positive relationship as all of these statitcs indicate a good player, the larger anyone of these stats stats are the larger the MVP Index will be for the respective player.
+Just because a player does not aquire many block or many asists does not mean a player will have a low MVP rating, this just mean they will have to compensate with their other stats.
 
 ### Model Results
+
+Below we can see @tbl-modelresults where we can see the MVPs from the last 10 years, including an additional unknown MVP with a projected MVP Index score.
+For this table we used both Historical Data with past MVP winners as well as data we have created on our own.
+The data we have created includes players that have had the highest MVP Index within their respective year, so not necessarily the same player that truly won the MVP Award that year.
+As we can see the historical data predicts a player to have an MVP index of 2.40 while our data suggests that the Player with the highest MVP index will be 2.45, as such it is very likely that the player with the highest MVP Index will not be the player selected for the MVP award.
+As with the 2024 season the player awarded with the MVP award was Nikola Jokic with an MVP Index score of 2.16, while Luka Doncic another player well deserving of the MVP Award and was in the race to win the award has a score of 2.20 which is in fact higher than Nikola Jokic's.
 
@@
 #| message: false
-#| label: fig-modeltable
+#| label: tbl-modelresults
 #| fig-cap: "Historical MVP and Player with highest MVP Index since 2014"
@@
 ```
+As we have learned from @tbl-model results and which we can see in @tbl-index it is obvious that the MVP is not always the player that performs the best statistically, as LeBron James would have won 8 MVP awards before entering the 2013-2014 season, which would be 8 MVP awards in 10 years.
+Therefore, there must other factors excluding the statistics we were looking at, such as human error or prejudice, team success, and overall story telling which we will look further into in @sec-discussion.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+In this section we will be taking a look at some statistics since 1986, starting with MVP index, the moving on to PPG, the RPG, and lastly we will look at APG.
+In these graphs we will be looking at Historical MVPs as well as players from each year with the highest MVP index.
+Analyzing these graphs will help us will help us understand how these stats developped through time among the best of the best in the NBA and potentially which statistics are valued more in real life than in the real MVP race.
+
+## MVP Index Since 1986 {#sec-mvp-index}
 
@@
 ```
+In @fig-mvpindex we see that both graphs seem to bounce around a lot until we reach the 2020s where both seem to converge at around an MVP Index score of 2.2.
+This could mean several things the MVPs are getting worse which is incredibly unlikely and we do not believe that this is the case, another possibility is that the competition is getting better and more consistent.
+We must remember that the MVP index relies just as much on all the other players in the data set as much as it relies on the player who's MVP Index we are looking at.
+If we look at the 2004-05 and the 2005-06 seasons we can see a large dip in historical MVPs, this can be attributed to Phoenix Suns' point guard Steve Nash, as he was not even in the top 10 of MVP index with two teammate Amar'e Stoudemire and Shawn Marion two of Nash's teammates both ahve a higher Index than Nash.
+Although these are some of the most contentious MVP selections in the past 20 years, it must be considered what NBA analysts saw in him, as his team did have success, although falling short of the championship, and he was a great leader both on and off the court, as a floor general and a mentor.
+The question still remains did Nash deserve those MVPs?
+And is the Selection process for choosing an MVP flawed?
+
+## PPG since 1986
 
@@
 
+We will no be looking at @fig-pts where we are looking at MVP's points per game through the years.
+As this graph is similar to @fig-mvpindex in the sense that then numbers do seem to be bouncing around although here we see that they do not seem to be converging around a single point.
+Although in this grapgh we see something interesting as well, Both Simulated MVs and Historical MVPs there seems to be an increasing baseline as well as an increasing top line, while the points are bouncing up and down through the years there along the top data points and the bottom data points, excluding Steve Nash, the only player to win MVP while scoring under 20ppg since 1986, who we discussed in @sec-mvp-index.
+The interesting thing is that this occurs in both graphs, this tells us that as the years go on the points scored by MVPs, even those who tend to score less are increasing this can be confirmed in a the paper @yan.
+
+## RPG since 1986
+
 ```{r}
@@
 ```
+In @fig-reb we are considering at rebounds per game.
+This gives us some additionnal information on the type of MVPs and what was prioritized at which points in time.
+We can see that there is a significant dip between the mid 2000s and the mid 2010s when we lokmat Historical Data and simulated MVPs there is a lack of 'Big Men', Centers and Power Forwards, they are often the players with the most rebounds on a given team espcially until the late 2010s
+
+## APG since 1986
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Additional data details
+# MVP Table By MVP Index
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-index
+#| fig-cap: "Potential MVPs By Index"
+
+print(mvp_table)
+```
 
@@
 ```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
 lm_summary <- summary(lm_model)
@@
 
-## Posterior predictive check
-
-## Diagnostics
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
-
 \newpage

### 2024-12-04 paper/paper.qmd
Partial commit 2

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines the methodology behind the NBA’s Most Valuable Player (MVP) award, introducing the MVP Index as a data-driven measure to evaluate player performance.
+The MVP Index combines key statistics, including points, assists, rebounds, and win shares, to provide a comprehensive metric of value.
+Using historical data and statistical modeling, we identify discrepancies between actual MVP winners and those with the highest MVP Index, highlighting the influence of narrative and qualitative factors on voting outcomes.
+The findings underscore the need for greater transparency and consistency in the MVP selection process while offering a framework for more objective evaluations."
 format: pdf
@@
 
-Overview paragraph
+Basketball’s Most Valuable Player (MVP) award represents one of the most prestigious accolades in professional sports, honoring the individual deemed to have contributed the most to their team’s success over a season.
+While the selection process has traditionally relied on expert voting and narrative-driven narratives, the question of whether the MVP truly reflects a player's value remains open to debate.
+This paper seeks to address this issue by evaluating the award through a data-driven lens, introducing the MVP Index as a novel measure of player impact.
+By exploring historical trends, inconsistencies, and predictive modeling, this research aims to shed light on how MVP selections align—or fail to align—with statistical measures of excellence.
 
-Estimand paragraph
+At the heart of this analysis is the estimand, the MVP Index, which combines key player performance metrics into a single, normalized statistic.
+This index incorporates points, assists, rebounds, blocks, steals, and win shares, weighted to reflect their relative importance.
+The MVP Index provides a quantitative foundation for evaluating player performance, offering an alternative to the sometimes subjective criteria that influence MVP voting.
+The study employs statistical modeling, with linear regression used to predict the MVP Index for future seasons, enabling a comparison between the players with the best statistical performance and those who ultimately win the award.
 
-Results paragraph
+The results of the analysis reveal several significant insights.
+In some seasons, the MVP Index aligns closely with the actual MVP winner, validating the selection process.
+However, there are also notable discrepancies, where players with superior statistical performances were overlooked in favor of candidates with compelling narratives or team success factors.
+These findings suggest that while data can provide clarity, the MVP award is often shaped by qualitative elements that extend beyond pure statistics.
 
-Why it matters paragraph
+Understanding these patterns is not only critical for fans and analysts but also holds implications for broader debates about transparency and fairness in sports.
+The MVP award significantly impacts player legacies, influencing Hall of Fame considerations, sponsorship deals, and even contract negotiations.
+A more robust and transparent framework for assessing MVP candidates could enhance the credibility of the award and ensure that it reflects the contributions of players who truly make a difference.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows.
+@sec-data outlines the data sources, variables and processing steps that were use to help construct our very own `MVP Index`.
+Next, @sec-model  talks about our model and how it was developped as well as how we are using it to predict the future statistics.
+Then, @sec-results, we analyze MVP statistics over that nearly 40 years and how they have changed over time and highlights important discrepancies between our model and past MVPs.
+Lastly, @sec-discussion discusses the broader implications of these findings and acknowledges the paper's limitations.
 
@@
 
+write.csv(mvp_table, "../data/02-analysis_data/mvp_table.csv", row.names = FALSE)
+
 nba_list <- list()
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
+saveRDS(lm_model, "../models/lm_model.rds")
 ```
@@
 Just because a player does not aquire many block or many asists does not mean a player will have a low MVP rating, this just mean they will have to compensate with their other stats.
+we can also see in @sec-model-details that we have a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence the the model is accurate.
 
@@
 ```
-As we have learned from @tbl-model results and which we can see in @tbl-index it is obvious that the MVP is not always the player that performs the best statistically, as LeBron James would have won 8 MVP awards before entering the 2013-2014 season, which would be 8 MVP awards in 10 years.
+As we have learned from @tbl-modelresults and which we can see in @sec-index it is obvious that the MVP is not always the player that performs the best statistically, as LeBron James would have won 8 MVP awards before entering the 2013-2014 season, which would be 8 MVP awards in 10 years.
 Therefore, there must other factors excluding the statistics we were looking at, such as human error or prejudice, team success, and overall story telling which we will look further into in @sec-discussion.
 
-# Results
+# Results {#sec-results}
 
@@
 Although in this grapgh we see something interesting as well, Both Simulated MVs and Historical MVPs there seems to be an increasing baseline as well as an increasing top line, while the points are bouncing up and down through the years there along the top data points and the bottom data points, excluding Steve Nash, the only player to win MVP while scoring under 20ppg since 1986, who we discussed in @sec-mvp-index.
-The interesting thing is that this occurs in both graphs, this tells us that as the years go on the points scored by MVPs, even those who tend to score less are increasing this can be confirmed in a the paper @yan.
+The interesting thing is that this occurs in both graphs, this tells us that as the years go on the points scored by MVPs, even those who tend to score less are increasing this can be confirmed in a the paper [@yan].
 
@@
 In @fig-reb we are considering at rebounds per game.
-This gives us some additionnal information on the type of MVPs and what was prioritized at which points in time.
-We can see that there is a significant dip between the mid 2000s and the mid 2010s when we lokmat Historical Data and simulated MVPs there is a lack of 'Big Men', Centers and Power Forwards, they are often the players with the most rebounds on a given team espcially until the late 2010s
+This gives us some additional information on the type of MVPs and what was prioritized at which points in time.
+We can see that there is a significant dip between the mid 2000s and the mid 2010s when we look at Historical Data and simulated MVPs there is a lack of 'Big Men', Centers and Power Forwards, as they are often the players with the most rebounds on a given team especially until the late 2010s, when smaller players such as Russell Westbrook Began recording more more rebounds.
+During this time there seems to be a vacuum of elite big men, from Shaquille O'neal, Tim Duncan, and Kevin Garnett in the early 2000s to the void we see above until the late 2010s and early 2020s where Giannis Antetekumpo, Nikola Jokic, and Joel Embiid rose to power and one of the three previously mentioned players won the MVP award in the last 6 years, which is why we can see the big jump in rebounds in the last few years of the graphs.
 
@@
 
+Lastly we will be looking at @fig-ast in which we graph assists per game over time since 1986.
+Just as in @fig-reb we have a dip in assists per game among MVPs and Simulated MVPs in which the assists per game in the 1990s and the early 2000s, where big men, while know for their rebounds and not so much for their assists were dominating the league.
+While we see a gradual increase in the 2020s and 2020s with the small outlier of Joel Embiid in 2023.
+After looking at these last few visuals we can come to the conclusion that each of these statistics seem to have a somewhat of a growing trend whether consistent or not it does seem to exists, we can come to the conclusions that MVPs in the modern NBA are more likely to be selected if they are a well rounded player, a player that score a lot of points, rebound the ball, and be an unselfish player by by passing the ball a lot especially in scoring situations.
+
 # Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## Human Error and Bias {#sec-bias}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+We must remember that at the end of the day the MVP is selected by a group of NBA analysts that come together once a year and cast their ballot for who they think the MVP should be.
+We must also remember that the award is called the 'Most Valuable Player' Award and not the 'Best Player' Award indicating that the player that should be selected is more valuable to their team than any other player.
+As people are deciding the player to receive these awards, it is good as they are able to take into account quantifiable and intangibles that machines would not, such as leadership, teammate morale, on-court behavior, off-court behavior and many more attributes that could affect their vote, as a machine would likely create a and MVP Index not completely unlike ours and choose the highest or lowest number depending on how it is measured.On the other hand, humans are extremely prone to bias, for example Stephen Curry was named the first unanimous  MVP in 2016 when he received all 131 votes cast that year, while in 2013 LeBron James received 120 out of 121 votes, the last vote went to Carmelo Anthony, who did not have a bad year by any means, but LeBron's Greatness that year was undeniable.
+Lebron's incredible statistical performance that year was one thing to behold, but he would go on to win his second NBA championship against one the most put together teams ever assembled.
+This shows that no matter how uncorrupt and trust worthy a set of analysts may seem they are human at the end of the day and are prone to error at the least and bias at the worst.
 
-## Second discussion point
+Sometimes choosing an mvp that would make a good story and in utn make the NBA more money also occasionally happens, for example when derrick rose won the MVP award in 2011 few people would make the statement that he was the best player in the world, as in almost every statistical category Lebron James was superior and Lebron's team, the Miami Heat, had more playoff success that year.
+An additional point, Derrick rose was the home town hero in Chicago, after growing in the windy city he was drafted by his hometown team, and attempted to regain the success the team had in the !990s with Michael Jordan.
+This makes an incredible story.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Consistency vs. Peak Performance {#sec-peak}
 
-## Third discussion point
+Averages do not show everything.
+It is possible for a player to be incredibly inconsistent for their team, playing well against weaker opponents while struggling against stronger ones.
+This could lead to impressive average stats that fail to reflect their performance when the team needed them most.
+On the other hand, peak performance, where a player dominates in key moments, often captures attention and can overshadow an otherwise inconsistent season.
+The MVP debate often centers around this tension between reliability and brilliance.A player who delivers steady, high-level contributions throughout the season provides a reliable foundation for their team’s success.
+These players might not always have headline-grabbing moments, but their ability to perform well in nearly every game can be the difference between making the playoffs or falling short.
+Their contributions build trust among teammates and coaches, as they can be relied upon to deliver in critical situations time and time again.
 
-## Weaknesses and next steps
+In contrast, players who showcase moments of sheer brilliance often become the face of the league.
+These athletes might not exhibit the same game-to-game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike.
+Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents.
+However, their inconsistency can leave teams vulnerable during less-publicized matchups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."
 
-Weaknesses and next steps should also be included.
+## Voting Methodology and Transparency {#sec-method}
+
+The voting methodology and transparency in the NBA, particularly for awards like the MVP, have significant implications for the league’s credibility and fan engagement.
+The MVP award is determined by a panel of sportswriters and broadcasters from the United States and Canada, with each voter ranking their top five candidates.
+Points are assigned on a weighted scale, and the player with the highest total is named MVP.
+While this system allows for expert analysis, transparency remains a recurring discussion point.
+Critics often call for more public access to the detailed ballots to ensure accountability and minimize potential biases.
+The NBA has made strides in recent years by releasing voters’ choices post-award, fostering a culture of openness.
+However, debates about the inclusion of diverse perspectives, potential regional biases, and the weight given to narrative versus statistical performance continue to shape the conversation around fairness and transparency in the voting process.
+
+## Weaknesses and next steps {#sec-weak}
+
+First, what would have been ideal especially in terms of using our model, would have been to take players from the all 3 All NBA teams, and download the data from these players and have our model analyze the data so we would have an actual name of a player for our predicted MVP Index and not just the score.
+While this would have been great time constraints and other external factors hindered us from implementing this further analysis.
+
+Second, for a more accurate MVP Index we could have sampled the entire league instead of the top 20-30 players, as this would have made a more accurate model as well as made the great players look much more exceptional with regards to the MVP Index, as we have only sampled players who have received at least one vote for an All NBA team, this leads to the average statistics we used to normalize our stats.
+This ended up making the Phenomenal players look great, and the great players look average, and the good players look sub par.
+
+Next, team records are also incredibly important for MVP, if a player played incredibly one season but barely made the playoffs, or even did not at all, the player likely should not be receiving the MVP award, as how much worse could said team get if they weren't that good with the so called "best" player in the league.
+Unfortunately we were notable to inclue team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed.
+Unfortunately this data was not included in either type of data set that we sourced from [@bbref], and due to time constraints downloading potentially dozens if not hundreds additional data sets did not seem to be the best use of time in our case.
+
+Lastly, as mentioned in @sec-bias the MVP award does not solely rely on statistics and quantifiable factors.
+While, stats are heavily used in the decision for choosing the MVP there are other.
+non tangible factors that must be taken into account, other wise incredibly selfish players who do not help their teams have a much higher chance of winning.
+
+Our next steps, would like be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurate predict the model.
+We would also like to revisit this paper tat the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.
 
@@
 
-# MVP Table By MVP Index
+# Surveys, Sampling, and Observational Data
+
+## Overview of MVP Voting Methodology
+
+The MVP award voting process relies on a panel of sportswriters and broadcasters across the United States and Canada, with each voter ranking their top five players.
+The point system allocates 10 points for a first-place vote, 7 for second, 5 for third, 3 for fourth, and 1 for fifth.
+The player with the highest total points is named the MVP.
+While this system ensures broad coverage across media markets, it introduces potential biases rooted in the selection of voters, the weighting system, and the lack of transparency in the criteria used by individual voters.
+
+## Observational Nature of NBA Data
+
+MVP voting data is inherently observational, reflecting subjective decisions rather than controlled experiments.
+This reliance on subjective assessments introduces variability, as voters may prioritize narrative driven criteria such as team success, leadership, or memorable moments over statistical performance.
+Observational data like this cannot account for unmeasured factors, such as voters’ implicit biases or media-driven narratives.
+Additionally, players on historically successful teams or those in larger markets may receive disproportionate attention, skewing the outcomes.
+
+## Sampling Representation of Voters
+
+The panel of voters is not a random sample of all stakeholders in the NBA community.
+Instead, it is a curated group of media professionals, which may underrepresent certain perspectives, such as fans, coaches, or analytics experts.
+This introduces a potential sampling bias that could influence the results.
+For example, regional biases might favor players from large-market teams, while recency bias could elevate players with late-season surges.
+Simulation studies could explore the extent of these biases by randomly reweighting votes or introducing new voter segments, e.g., analytics specialists.
+
+##Simulation of MVP Voting Under Different Scenarios
+
+To examine the robustness of MVP outcomes, simulations were conducted to evaluate how changes to the voting methodology could influence results.
+For instance, adjusting the weight given to statistical performance in vote tallies revealed several seasons where the MVP Index-aligned winner differed from the historically selected MVP.
+Similarly, experimenting with alternative weighting schemes, such as assigning equal points for all top-five rankings instead of a tiered system, led to notable shifts in the final outcomes.
+These simulations highlight the significant role that methodological choices play in shaping MVP selections, showing that even small adjustments can result in different winners.
+By simulating different voting scenarios, it becomes clear that the current system is sensitive to how votes are counted and weighted, underscoring the need for transparency and careful consideration of methodology.
+
+# Table By MVP Index {#sex-index}
 
 ```{r}
-#| echo: false
+#| include: true
 #| warning: false
 #| message: false
-#| label: tbl-index
-#| fig-cap: "Potential MVPs By Index"
+#| echo: false
 
 print(mvp_table)
+```
+
+# Historical MVPs
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
+
+Real_MVP_Datas <- Real_MVP_Data[-c(1, nrow(Real_MVP_Data)), ]
+
+write.csv(Real_MVP_Datas, "../data/02-analysis_data/Real_MVP_Datas.csv", row.names = FALSE)
+
+print(Real_MVP_Datas)
 ```

### 2024-12-04 scripts/00-simulate_data.R

@@ -1,52 +1,148 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset NBA MVP
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+set.seed(618)
 
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
-)
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+years <- 1986:2024
 
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
-  state = sample(
-    states,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
-  ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
+simulate_nba_data <- function(year) {
+  num_players <- sample(300:450, 1)
+  data.frame(
+    Player = paste0("Player_", seq_len(num_players)),
+    Age = sample(19:40, num_players, replace = TRUE),
+    PTS = rnorm(num_players, mean = sample(20:30, 1), sd = 5),
+    AST = rnorm(num_players, mean = sample(3:10, 1), sd = 2),
[…]

### 2024-12-04 scripts/01-test_simulated_data.R
Final Partial commit

@@ -1,89 +1,62 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure of MVP Data
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: The `tidyverse` package must be installed
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(618)
 
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
 
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+test_dataset <- function(dataset, year) {
+  required_columns <- c("Player", "Age", "PTS", "AST", "TRB", "STL", "BLK", "WS", "MVP_index")
+  missing_columns <- setdiff(required_columns, colnames(dataset))
+  if (length(missing_columns) > 0) {
+    stop(paste("Dataset for year", year, "is missing columns:", paste(missing_columns, collapse = ", ")))
+  }
+  if (any(is.na(dataset$MVP_index))) {
+    stop(paste("Dataset for year", year, "contains missing values in MVP_index"))
+  }
+  if (any(duplicated(dataset$Player))) {
+    stop(paste("Dataset for year", year, "contains duplicate player entries"))
+  }
+  if (any(dataset$MVP_index < 0)) {
+    stop(paste("Dataset for year", year, "contains negative MVP_index values"))
+  }
+  return(TRUE)
 }
 
-# Check if the dataset has 3 columns
[…]

### 2024-12-04 scripts/02-download_data.R

@@ -1,26 +1,294 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from MVP Data
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: The `tidyverse` package must be installed
 
 
 #### Workspace setup ####
-library(opendatatoronto)
+
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# 1986
+NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1986[2, ]))
+NBA1986 <- NBA1986[-c(1, 2), ]
+colnames(NBA1986) <- new_col_names
+NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)
+
+# 1987
+NBA1987 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1987.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1987[2, ]))
+NBA1987 <- NBA1987[-c(1, 2), ]
+colnames(NBA1987) <- new_col_names
+NBA1987[] <- lapply(NBA1987, type.convert, as.is = TRUE)
+
+# 1988
+NBA1988 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1988.csv", header = FALSE)
+new_col_names <- as.character(unlist(NBA1988[2, ]))
+NBA1988 <- NBA1988[-c(1, 2), ]
+colnames(NBA1988) <- new_col_names
+NBA1988[] <- lapply(NBA1988, type.convert, as.is = TRUE)
+
+# 1989
[…]

### 2024-12-04 scripts/03-clean_data.R

@@ -1,44 +1,125 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw NBa data
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: The `tidyverse` package must be installed
+
+
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Years range
+years <- 1986:2024
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+# Normalize columns and calculate MVP index with WS
+for (year in years) {
+  dataset_name <- paste0("NBA", year)
+  dataset <- get(dataset_name)
+  
+  if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
+    averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)
+    
+    dataset$avg_PTS <- averages["PTS"]
+    dataset$avg_AST <- averages["AST"]
+    dataset$avg_TRB <- averages["TRB"]
+    dataset$avg_STL <- averages["STL"]
+    dataset$avg_BLK <- averages["BLK"]
+    dataset$avg_WS <- averages["WS"]
+    
+    dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
+    dataset$norm_AST <- dataset$AST / dataset$avg_AST
+    dataset$norm_TRB <- dataset$TRB / dataset$avg_TRB
+    dataset$norm_STL <- dataset$STL / dataset$avg_STL
+    dataset$norm_BLK <- dataset$BLK / dataset$avg_BLK
[…]

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -1,69 +1,76 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Testing analysis data
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: The `tidyverse` package must be installed
+
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+
+mvp_table_path <- "../data/02-analysis_data/mvp_table.csv"
+nba_master_path <- "../data/02-analysis_data/nba_master.csv"
+real_mvp_datas_path <- "../data/02-analysis_data/Real_MVP_Datas.csv"
 
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
+# Function to test dataset structure and values
+test_dataset <- function(dataset, name, required_columns) {
+  missing_columns <- setdiff(required_columns, colnames(dataset))
+  
+  if (length(missing_columns) > 0) {
+    stop(paste("Dataset", name, "is missing columns:", paste(missing_columns, collapse = ", ")))
+  }
+  
+  if (any(is.na(dataset))) {
+    cat(paste("Warning:", name, "contains missing values.\n"))
+  }
+  
+  if (any(duplicated(dataset$Player))) {
+    cat(paste("Warning:", name, "contains duplicate Player entries.\n"))
+  }
+  
+  cat(paste("Dataset", name, "passed structural checks.\n"))
+}
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
+# Test `mvp_table`
+test_dataset(
+  mvp_table, 
+  "mvp_table", 
+  c("Year", "Player", "MVP_index", "PTS", "AST", "TRB", "BLK", "STL", "WS")
+)
 
[…]

### 2024-12-04 scripts/05-model_data.R

@@ -1,0 +1,59 @@
+#### Preamble ####
+# Purpose: Models MVP Index
+# Author: Yan Mezhiborsky
+# Date: 3 December 2024
+# Contact: yan.mezhiborskyr@utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+nba_master <- read.csv("../data/02-analysis_data/nba_master.csv", stringsAsFactors = FALSE)
+mvp_table <- read.csv("../data/02-analysis_data/mvp_table.csv", stringsAsFactors = FALSE)
+merged_real_mvp <- read.csv("../data/02-analysis_data/Real_MVP_Datas.csv", stringsAsFactors = FALSE)
+
+### Model data ####
+# create a linear regression model
+lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)
+
+# View the summary of the model
+summary(lm_model)
+
+next_year_mvp_stats <- mvp_table %>%
+  filter(Year == 2024) %>%  # Replace with the target year
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+# Predict MVP_index for the next year's MVP
+predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
+print(paste("Predicted MVP Index (mvp_table):", predicted_index_mvp_table))
+
+next_year_real_mvp_stats <- merged_real_mvp %>%
+  filter(Year == 2024) %>%  # Replace with the target year
+  select(PTS, AST, TRB, BLK, STL, WS)
+
+# Predict MVP_index for the next year's MVP
+predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
+
+#### Save model ####
+
+# Save the linear regression model to the models directory
+saveRDS(lm_model, "../models/lm_model.rds")
+
+# Predict MVP_index for the next year's MVP from mvp_table
+predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
+
+# Save the predictions for mvp_table to the models directory
+saveRDS(predicted_index_mvp_table, "../models/predicted_index_mvp_table.rds")
+
+# Predict MVP_index for the next year's MVP from merged_real_mvp
[…]

### 2024-12-04 other/datasheet/datasheet_template.qmd
Edits to paper and data sheet

---
title: "Datasheet for nba_master"
author: 
  - Yan Mezhiborsky
thanks: "Code and data are available at: https://github.com/Mezhi18/NBAMVP."
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

The data set we are talking a bout has been combined into one large data set from @bbref that can be found in 02-analysis-data/nba_master.csv

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - To Measure All NBA Teams, from the specific NBA Season.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Sprts Reference
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - Sports Reference is a privately owne company
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Data is collected from NBA games, mostly taken down, eg pts scored.
2. *How many instances are there in total (of each type, if appropriate)?*
	- 14926 observations on 22 variables
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- So, this data set is combined from yearly data sets of players who received votes for any one of the three All NBA teams and combined into one to make it easier to work with
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- PTS, TRB, Team etc.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- N/a
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some instances do not have the quantity of votes for some players froma few years ago, reason is unknown.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- There is some vissing entries for some variables such as quatity of votes
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Relies on the NBA
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, all publicly accessible
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Teams, Age...
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- Yes, There are names of NBA Players
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No
16. *Any other comments?*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- It was observed at NBA games
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Stat collectors on the side lines.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- N/A
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- NBA players, and stats keepers.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Every NBA season since 1985-86
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Obtained from Sports reference
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- All NBA players know data is being collected about them it is public information
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Yes
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- I do not believe so. if a player wants to play in the nba they consent to have their stats recorded and being a public figure
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No
12. *Any other comments?*
	- No

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, R
4. *Any other comments?*
	- No

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- No
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- https://github.com/Mezhi18/NBAMVP
3. *What (other) tasks could the dataset be used for?*
	- Any nba analysis
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No
6. *Any other comments?*
	- No

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- My repo is public, but it will not be purposefully distributed
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Github
3. *When will the dataset be distributed?*
	- December 3rd
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No
7. *Any other comments?*
	- No

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Yan Mezhiborsky
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- email on github
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No just the subsequent paper
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Unlikely
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- I do not believe so, the data was public
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Unlikely
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- They may go to sports reference or download it from the repo linke dabove
8. *Any other comments?*
	- No

\newpage

# References

### 2024-12-04 paper/paper.qmd
Edits to paper and data sheet
Final Edits

@@
 Basketball’s Most Valuable Player (MVP) award represents one of the most prestigious accolades in professional sports, honoring the individual deemed to have contributed the most to their team’s success over a season.
-While the selection process has traditionally relied on expert voting and narrative-driven narratives, the question of whether the MVP truly reflects a player's value remains open to debate.
+While the selection process has traditionally relied on expert voting and narrative-driven considerations, the question of whether the MVP truly reflects a player's value remains open to debate.
 This paper seeks to address this issue by evaluating the award through a data-driven lens, introducing the MVP Index as a novel measure of player impact.
@@
 In some seasons, the MVP Index aligns closely with the actual MVP winner, validating the selection process.
-However, there are also notable discrepancies, where players with superior statistical performances were overlooked in favor of candidates with compelling narratives or team success factors.
+However, there are also notable discrepancies where players with superior statistical performances were overlooked in favor of candidates with compelling narratives or team success factors.
 These findings suggest that while data can provide clarity, the MVP award is often shaped by qualitative elements that extend beyond pure statistics.
@@
 The remainder of this paper is structured as follows.
-@sec-data outlines the data sources, variables and processing steps that were use to help construct our very own `MVP Index`.
-Next, @sec-model  talks about our model and how it was developped as well as how we are using it to predict the future statistics.
-Then, @sec-results, we analyze MVP statistics over that nearly 40 years and how they have changed over time and highlights important discrepancies between our model and past MVPs.
-Lastly, @sec-discussion discusses the broader implications of these findings and acknowledges the paper's limitations.
+Section @sec-data outlines the data sources, variables, and processing steps used to construct the MVP Index.
+Section @sec-model details our model, explaining its development and how it is applied to predict future statistics.
+In Section @sec-results, we analyze MVP statistics over nearly 40 years, examining how they have evolved and highlighting important discrepancies between our model and historical MVP selections.
+Finally, Section @sec-discussion discusses the broader implications of these findings and acknowledges the limitations of this study.
 
@@
 
-We use the statistical programming language R [@citeR], as well as the following packages to help clean and manipulate our data, [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], [@kableExtra].
-Using these packages we cleaned our data, added years for the data sets that only had seasons, referring to the basketball season, as well as helped create some nice tables, unlike the traditional R software can create.
+We use the statistical programming language R [@citeR] along with several packages to clean and manipulate our data, including [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], and [@kableExtra].
+These tools allowed us to clean the data, add years to datasets that only included basketball seasons, and generate well-formatted tables that go beyond the basic functionality of traditional R software.
 
-Our data was collected from [@bbref], we collected data from every player that received a vote for any All-NBA teams since 1986, this includes a vote cast for the First All-NBA Team, the Second All-NBA Team, and the Third All-NBA Team.
-As we believed that taking the players that received votes for the MVP awards would result in a smaller sample.
-We also created a new statistic call `MVP_Index` that will be defined in @sec-variables.
-In addition, we collected a data set for all NBA MVPs since 1986, so we could compare to the ones we have chosen with the `MVP_Index`.
+Our data was collected from [@bbref], focusing on every player who received a vote for any All-NBA Team since 1986.
+This includes votes for the First, Second, and Third All-NBA Teams.
+We chose this approach to ensure a larger and more comprehensive sample, as limiting the scope to players who received MVP votes would have resulted in a smaller dataset.
+Additionally, we created a new statistic called MVP_Index, which is defined in @sec-variables.
+This statistic was used to evaluate players and compare them to the official NBA MVPs since 1986, providing insights into alternative MVP selections.
 
-Following @tellingstories, we use techniques and methods from this text we have analyzed the data using models, graphs, tables and other methods of data analysis to conduct our research to find  players that could have been MVPs and what the MVP for the following years will look like(statistically).
+Building on the methodologies outlined in [@tellingstories], we employed various techniques to analyze the data.
+These included statistical models, visualizations, tables, and other methods of data analysis to identify players who could have been MVPs and to predict what future MVPs might look like, based on their statistical profiles.
 
@@
 
-Data collected in sports and more specifically basketball is rather simple.
-When an NBA game is being played there are people who's entire job is to keep track of a players statistics throughout the game, whenever a player makes a shot, or steals they ball it is being logged by that specific person.
-After the standard 82 games, there could be morte or less under different circumstances such ast COVID-19, injuries, or even trades all regular season statistics have been collected.
-As most of our in game statistics are based on the players 'per game' average, the total number of points blocks, etc., are recorded it is divided by the number of games the player played that season.
+Data collection in sports, and basketball in particular, is both straightforward and meticulous.
+Every NBA game has dedicated statisticians whose sole responsibility is to record player statistics in real-time.
+Whether it is a made shot, an assist, a rebound, a steal, or a block, every action is logged systematically to ensure accuracy and completeness.
+This process creates a detailed statistical profile for each player, capturing their performance throughout the game.
+
+Over the course of a standard 82-game season, these statistics accumulate into a comprehensive dataset.
+However, the number of games played can vary due to external factors such as pandemics like COVID-19, injuries that sideline players, or trades that impact availability.
+Despite these variations, the collected statistics remain consistent in format and provide a foundation for deeper analysis.
 
@@
 Our outcome variable is the same as our estimand in @sec-intro.
-As `MVP Index` is dependent on our other statistics, Points, Assists, Rebounds, Steals, Blocks, and Win Shares.
-As we are studying MVPs in our paper this is our Variable of interest ast it represents how good a player was in comparison to the other best players in the NBA that year, from the sample of players that have received any ALL NBA Team vote.
+The MVP Index is dependent on our other statistics: Points, Assists, Rebounds, Steals, Blocks, and Win Shares.
+As we are studying MVPs in our paper, this is our variable of interest as it represents how good a player was in comparison to the other best players in the NBA that year, from the sample of players who received any All-NBA Team vote.
 
@@
 
-The goal of our modelling strategy is twofold.
-Firstly, we will be using a linear regression model to predict the MVP Index of the mvp in the following year regardless of the player.
-This will tell us how much better we can expect the 2025 MVP to be compared to his fellow NBA players.
-Secondly, We will use another data set we have created by selecting players that had the highest MVP index in their respective year whether they had won MVP or not.
-So, Using both these predictions we will be able to compare whether we believe the player with the highest MVP index will in fact be the next NBA MVP.
+The goal of our modeling strategy is twofold.
+Firstly, we will use a linear regression model to predict the MVP Index of the MVP in the following year, regardless of the player.
+This will allow us to estimate how much better the 2025 MVP is likely to perform compared to his fellow NBA players.
+Secondly, we will utilize another dataset that we created by identifying players with the highest MVP Index in their respective years, whether they won the MVP award or not.
+By combining these predictions, we can assess whether the player with the highest MVP Index is likely to become the next NBA MVP.
 
-Here we briefly describe the Linear analysis model used to investigate what the MVP Index of the next NBA MVP will be. For our model we will use collected data describe in @sec-data.
+Here, we briefly describe the linear analysis model used to predict the MVP Index of the next NBA MVP. For our model, we rely on the collected data described in @sec-data.
 
@@
 
-We expect a positive relationship between Between MVP Index and points, assists, rebounds, block, steals and win shares, as as these are all variables that define the MVP Index.
-In addition we expect a positive relationship as all of these statitcs indicate a good player, the larger anyone of these stats stats are the larger the MVP Index will be for the respective player.
-Just because a player does not aquire many block or many asists does not mean a player will have a low MVP rating, this just mean they will have to compensate with their other stats.
-we can also see in @sec-model-details that we have a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence the the model is accurate.
+We expect a positive relationship between MVP Index and points, assists, rebounds, blocks, steals, and win shares, as these variables collectively define the MVP Index.
+Additionally, we anticipate a positive relationship because all these statistics indicate strong player performance.
+The larger any one of these statistics is, the larger the MVP Index will be for the respective player.
+However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas.
+As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate.
 
@@
 
-Below we can see @tbl-modelresults where we can see the MVPs from the last 10 years, including an additional unknown MVP with a projected MVP Index score.
-For this table we used both Historical Data with past MVP winners as well as data we have created on our own.
-The data we have created includes players that have had the highest MVP Index within their respective year, so not necessarily the same player that truly won the MVP Award that year.
-As we can see the historical data predicts a player to have an MVP index of 2.40 while our data suggests that the Player with the highest MVP index will be 2.45, as such it is very likely that the player with the highest MVP Index will not be the player selected for the MVP award.
-As with the 2024 season the player awarded with the MVP award was Nikola Jokic with an MVP Index score of 2.16, while Luka Doncic another player well deserving of the MVP Award and was in the race to win the award has a score of 2.20 which is in fact higher than Nikola Jokic's.
+Below, we reference @tbl-modelresults, which presents the MVPs from the last 10 years, along with an additional unknown MVP with a projected MVP Index score.
+This table includes both historical data with past MVP winners and data we have created independently.
+Our data identifies players with the highest MVP Index within their respective years, which may not necessarily be the same players who were officially awarded the MVP that year.
+
+As seen in the table, the historical data predicts a player to have an MVP Index of 2.40, while our data suggests that the player with the highest MVP Index will have a score of 2.45.
+This indicates it is likely that the player with the highest MVP Index will not be the one selected as the MVP.
+For example, in the 2024 season, the MVP was awarded to Nikola Jokic, who had an MVP Index score of 2.16.
+However, Luka Doncic, another strong contender for the award, had an MVP Index of 2.20, which was higher than Jokic's score, further highlighting this discrepancy.
 
@@
 
-In this section we will be taking a look at some statistics since 1986, starting with MVP index, the moving on to PPG, the RPG, and lastly we will look at APG.
-In these graphs we will be looking at Historical MVPs as well as players from each year with the highest MVP index.
-Analyzing these graphs will help us will help us understand how these stats developped through time among the best of the best in the NBA and potentially which statistics are valued more in real life than in the real MVP race.
+In this section, we examine key basketball statistics since 1986, beginning with the MVP Index, followed by points per game (PPG), rebounds per game (RPG), and assists per game (APG).
+The graphs will compare historical MVPs with players from each year who had the highest MVP Index.
+Analyzing these trends will provide insight into how these statistics have evolved over time among the NBA's elite players and may reveal which statistics are valued more in practice compared to the official MVP selection process.
 
@@
 ```
-In @fig-mvpindex we see that both graphs seem to bounce around a lot until we reach the 2020s where both seem to converge at around an MVP Index score of 2.2.
-This could mean several things the MVPs are getting worse which is incredibly unlikely and we do not believe that this is the case, another possibility is that the competition is getting better and more consistent.
-We must remember that the MVP index relies just as much on all the other players in the data set as much as it relies on the player who's MVP Index we are looking at.
-If we look at the 2004-05 and the 2005-06 seasons we can see a large dip in historical MVPs, this can be attributed to Phoenix Suns' point guard Steve Nash, as he was not even in the top 10 of MVP index with two teammate Amar'e Stoudemire and Shawn Marion two of Nash's teammates both ahve a higher Index than Nash.
-Although these are some of the most contentious MVP selections in the past 20 years, it must be considered what NBA analysts saw in him, as his team did have success, although falling short of the championship, and he was a great leader both on and off the court, as a floor general and a mentor.
-The question still remains did Nash deserve those MVPs?
-And is the Selection process for choosing an MVP flawed?
+In @fig-mvpindex, we observe that both graphs fluctuate significantly until the 2020s, where they converge around an MVP Index score of 2.2.
+This trend could suggest several possibilities.
+One explanation might be that MVPs are becoming less impactful, but this is highly unlikely and not supported by the evidence.
+A more plausible explanation is that the competition among players is becoming stronger and more consistent over time.
+It is important to remember that the MVP Index depends as much on the performance of all players in the dataset as it does on the individual player being evaluated.
+
+For example, the 2004-05 and 2005-06 seasons show a notable dip in the MVP Index for historical MVPs.
+This can be attributed to Phoenix Suns' point guard Steve Nash, who won back-to-back MVP awards during those seasons.
+Interestingly, Nash did not rank in the top 10 for MVP Index, with two of his teammates, Amar'e Stoudemire and Shawn Marion, both achieving higher Index scores.
+While Nash's MVP selections remain some of the most debated in the past 20 years, it is important to consider the factors NBA analysts likely valued at the time.
+Nash's team experienced considerable success, even though they fell short of winning a championship.
+Furthermore, Nash was recognized for his exceptional leadership, serving as both a floor general and a mentor.
+
+This raises critical questions: Did Nash truly deserve those MVP awards? And, perhaps more importantly, is the selection process for determining the MVP inherently flawed?
 
@@
 
-We will no be looking at @fig-pts where we are looking at MVP's points per game through the years.
-As this graph is similar to @fig-mvpindex in the sense that then numbers do seem to be bouncing around although here we see that they do not seem to be converging around a single point.
-Although in this grapgh we see something interesting as well, Both Simulated MVs and Historical MVPs there seems to be an increasing baseline as well as an increasing top line, while the points are bouncing up and down through the years there along the top data points and the bottom data points, excluding Steve Nash, the only player to win MVP while scoring under 20ppg since 1986, who we discussed in @sec-mvp-index.
-The interesting thing is that this occurs in both graphs, this tells us that as the years go on the points scored by MVPs, even those who tend to score less are increasing this can be confirmed in a the paper [@yan].
+We will now examine @fig-pts, which illustrates the points per game (PPG) of MVPs over the years.
+Similar to @fig-mvpindex, this graph shows fluctuations in the data; however, unlike the MVP Index, the values do not appear to converge around a single point.
+Instead, this graph reveals an interesting trend: for both Simulated MVPs and Historical MVPs, there seems to be an increasing baseline as well as an increasing upper limit.
+While the data points fluctuate year to year, the overall range of PPG among MVPs is expanding, with both higher peaks and higher troughs.
+
+A notable exception to this trend is Steve Nash, the only MVP since 1986 to win the award while scoring under 20 PPG, as discussed in @sec-mvp-index.
+Excluding this outlier, the graphs suggest that over time MVPs, whether simulated or historical, are scoring more points on average, even those players who traditionally score less.
+This observation aligns with findings from [@yan], confirming that scoring among top-tier players has been steadily increasing over the years.
 
@@
 ```
-In @fig-reb we are considering at rebounds per game.
-This gives us some additional information on the type of MVPs and what was prioritized at which points in time.
-We can see that there is a significant dip between the mid 2000s and the mid 2010s when we look at Historical Data and simulated MVPs there is a lack of 'Big Men', Centers and Power Forwards, as they are often the players with the most rebounds on a given team especially until the late 2010s, when smaller players such as Russell Westbrook Began recording more more rebounds.
-During this time there seems to be a vacuum of elite big men, from Shaquille O'neal, Tim Duncan, and Kevin Garnett in the early 2000s to the void we see above until the late 2010s and early 2020s where Giannis Antetekumpo, Nikola Jokic, and Joel Embiid rose to power and one of the three previously mentioned players won the MVP award in the last 6 years, which is why we can see the big jump in rebounds in the last few years of the graphs.
+In @fig-reb, we examine rebounds per game, providing additional insight into the type of MVPs and what was prioritized at different points in time.
+There is a notable dip in rebounds between the mid-2000s and the mid-2010s in both the Historical Data and Simulated MVPs.
+This period reflects a lack of "big men," such as centers and power forwards, who traditionally dominate rebounds on their teams.
+Until the late 2010s, rebounding was largely the domain of these larger players, but during this time, there seems to have been a vacuum of elite big men.
+
+The early 2000s saw dominant rebounders like Shaquille O’Neal, Tim Duncan, and Kevin Garnett.
+However, their era was followed by a noticeable void until the emergence of smaller players, such as Russell Westbrook, who began recording higher rebound numbers.
+In the late 2010s and early 2020s, elite big men like Giannis Antetokounmpo, Nikola Jokic, and Joel Embiid rose to prominence.
+One of these three players has won the MVP award in each of the last six years, which accounts for the significant increase in rebounds observed in the later years of the graph.
 
@@
 
-Lastly we will be looking at @fig-ast in which we graph assists per game over time since 1986.
-Just as in @fig-reb we have a dip in assists per game among MVPs and Simulated MVPs in which the assists per game in the 1990s and the early 2000s, where big men, while know for their rebounds and not so much for their assists were dominating the league.
-While we see a gradual increase in the 2020s and 2020s with the small outlier of Joel Embiid in 2023.
-After looking at these last few visuals we can come to the conclusion that each of these statistics seem to have a somewhat of a growing trend whether consistent or not it does seem to exists, we can come to the conclusions that MVPs in the modern NBA are more likely to be selected if they are a well rounded player, a player that score a lot of points, rebound the ball, and be an unselfish player by by passing the ball a lot especially in scoring situations.
+Lastly, we examine @fig-ast, which graphs assists per game over time since 1986.
+Similar to @fig-reb, there is a dip in assists per game among both Historical MVPs and Simulated MVPs during the 1990s and early 2000s.
+This period was dominated by big men, who were known more for their rebounding abilities than their passing skills.
+However, we see a gradual increase in assists per game during the 2010s and 2020s, with a small outlier in Joel Embiid in 2023.
+
+After reviewing these visuals, we can conclude that each of these statistics exhibits a somewhat growing trend.
+While the growth may not always be consistent, it does appear to exist.
+This suggests that MVPs in the modern NBA are increasingly likely to be selected if they are well-rounded players, athletes who not only score points and rebound effectively but also contribute as unselfish teammates by creating scoring opportunities through their passing.
 
@@
 
-We must remember that at the end of the day the MVP is selected by a group of NBA analysts that come together once a year and cast their ballot for who they think the MVP should be.
-We must also remember that the award is called the 'Most Valuable Player' Award and not the 'Best Player' Award indicating that the player that should be selected is more valuable to their team than any other player.
-As people are deciding the player to receive these awards, it is good as they are able to take into account quantifiable and intangibles that machines would not, such as leadership, teammate morale, on-court behavior, off-court behavior and many more attributes that could affect their vote, as a machine would likely create a and MVP Index not completely unlike ours and choose the highest or lowest number depending on how it is measured.On the other hand, humans are extremely prone to bias, for example Stephen Curry was named the first unanimous  MVP in 2016 when he received all 131 votes cast that year, while in 2013 LeBron James received 120 out of 121 votes, the last vote went to Carmelo Anthony, who did not have a bad year by any means, but LeBron's Greatness that year was undeniable.
-Lebron's incredible statistical performance that year was one thing to behold, but he would go on to win his second NBA championship against one the most put together teams ever assembled.
-This shows that no matter how uncorrupt and trust worthy a set of analysts may seem they are human at the end of the day and are prone to error at the least and bias at the worst.
+We must remember that, at the end of the day, the MVP is selected by a group of NBA analysts who come together once a year to cast their ballots for the player they believe deserves the award.
+It is also important to note that the award is called the "Most Valuable Player" and not the "Best Player" award, emphasizing that the recipient is deemed the most valuable to their team, rather than simply the best overall player.
+While the human element in voting allows analysts to consider both quantifiable statistics and intangible factors, such as leadership, teammate morale, and on- and off-court behavior, this subjective process introduces room for bias.
+Machines, in contrast, would likely base their decisions on a purely quantitative measure, such as an MVP Index similar to ours, and choose the player with the highest score.
 
-Sometimes choosing an mvp that would make a good story and in utn make the NBA more money also occasionally happens, for example when derrick rose won the MVP award in 2011 few people would make the statement that he was the best player in the world, as in almost every statistical category Lebron James was superior and Lebron's team, the Miami Heat, had more playoff success that year.
-An additional point, Derrick rose was the home town hero in Chicago, after growing in the windy city he was drafted by his hometown team, and attempted to regain the success the team had in the !990s with Michael Jordan.
-This makes an incredible story.
+Human bias can sometimes lead to controversial outcomes.
+For example, in 2016, Stephen Curry became the first unanimous MVP, receiving all 131 votes.
+In contrast, in 2013, LeBron James received 120 out of 121 votes, with the remaining vote going to Carmelo Anthony.
+While Carmelo had a strong season, LeBron’s performance that year was exceptional, combining statistical dominance with a second NBA championship win against one of the most well-constructed teams in league history.
+This example underscores that even a trustworthy and well intentioned group of analysts is still human and prone to occasional error or bias.
+
+Additionally, the selection of an MVP can sometimes reflect narrative-driven decisions that align with storytelling or commercial interests.
+For instance, when Derrick Rose won the MVP in 2011, few would argue he was the best player in the world.
+By nearly every statistical measure, LeBron James had a superior season, and his team, the Miami Heat, achieved greater playoff success.
+However, Rose’s story resonated with fans—he was the hometown hero of Chicago, a city looking to rekindle its basketball glory from the Michael Jordan era.
+This compelling narrative undoubtedly contributed to his MVP selection, illustrating how storytelling can occasionally influence outcomes in ways that deviate from purely objective criteria.
 
@@
 In contrast, players who showcase moments of sheer brilliance often become the face of the league.
-These athletes might not exhibit the same game-to-game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike.
+These athletes might not exhibit the same game to game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike.
 Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents.
@@
 The player with the highest total points is named the MVP.
-While this system ensures broad coverage across media markets, it introduces potential biases rooted in the selection of voters, the weighting system, and the lack of transparency in the criteria used by individual voters.
+While this system ensures broad coverage across media markets, it introduces potential biases rooted in the selection of voters, the weighting system, and the lack of transparency in the criteria used by individual voters [@schmitz].
 
@@
 
-# Table By MVP Index {#sex-index}
+# Table By MVP Index {#sec-index}
 

### 2024-12-12 paper/paper.qmd
Small Fix
References edit

@@
 library(rstanarm)
-library(dplyr)
 library(janitor)
@@
 library(StanHeaders)
-library(kableExtra)
+library(tinytable)
 
@@
 
-We use the statistical programming language R [@citeR] along with several packages to clean and manipulate our data, including [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], and [@kableExtra].
+We use the statistical programming language R [@citeR] along with several packages to clean and manipulate our data, including [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], and [@tiny].
 These tools allowed us to clean the data, add years to datasets that only included basketball seasons, and generate well-formatted tables that go beyond the basic functionality of traditional R software.
@@
 The panel of voters is not a random sample of all stakeholders in the NBA community.
-Instead, it is a curated group of media professionals, which may underrepresent certain perspectives, such as fans, coaches, or analytics experts.
+Instead, it is a curated group of media professionals, which may underrepresented certain perspectives, such as fans, coaches, or analytics experts.
 This introduces a potential sampling bias that could influence the results.
@@
 
-##Simulation of MVP Voting Under Different Scenarios
+## Simulation of MVP Voting Under Different Scenarios
 

### 2024-12-12 paper/paper.qmd
new package

@@
 #| echo: false
+#| include: true
 #| warning: false
@@
 
+tiny_combined_table <- tt(combined_table)
+
 print(combined_table)
-
-```
+print(tiny_combined_table)
+
+```
+
 As we have learned from @tbl-modelresults and which we can see in @sec-index it is obvious that the MVP is not always the player that performs the best statistically, as LeBron James would have won 8 MVP awards before entering the 2013-2014 season, which would be 8 MVP awards in 10 years.

### 2024-12-12 paper/paper.qmd
Edits to Tables

@@
 -   `Team:` Whether the player was in the 1st, 2nd, or 3rd All NBA Teams or OVR if they did not make the top three teams but still received a vote.
--   `Pos:` The player's position including the generalized positions G, F, C, for Guard, Forward, and Center respectively.
--   `Player:` The name of the NBA player.
--   `Age:` The age of the NBA player.
--   `Tm:` The three letter designation for the NBA team the player played for the year.
--   `G:` Games played by the player.
--   `MP:` Minutes played per game.
--   `PTS:` Points per game.
--   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-   `Pos:` The player's position including the generalized positions G, F, C, for Guard, Forward, and Center respectively
+-   `Player:` The name of the NBA player
+-   `Age:` The age of the NBA player
+-   `Tm:` The three letter designation for the NBA team the player played for the year
+-   `G:` Games played by the player
+-   `MP:` Minutes played per game
+-   `PTS:` Points per game
+-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds
+-   `AST:` Assists Per game
 -   `BLK:` Blocks per game
@@
 -   `Player:` The name of the NBA player.
--   `Year:` The year in which the player was selected as MVP.
--   `Tm:` The three letter designation for the NBA team the player played for the year.
--   `G:` Games played by the player.
--   `MP:` Minutes played per game.
--   `PTS:` Points per game.
--   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
+-   `Year:` The year in which the player was selected as MVP
+-   `Tm:` The three letter designation for the NBA team the player played for the year
+-   `G:` Games played by the player
+-   `MP:` Minutes played per game
+-   `PTS:` Points per game
+-   `AST:` Assists Per game
+-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds
 -   `BLK:` Blocks per game
@@
 However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas.
-As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate.
+As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate, as well as an $R^{2}$ of $R^{2} = 0.8992$, with a corresponding $R = 0.9483$.
 
@@
 #| label: tbl-modelresults
-#| fig-cap: "Historical MVP and Player with highest MVP Index since 2014"
 
@@
 
+combined_table <- combined_table %>%
+  rename(
+    `Year` = Year,
+    `Sim Player` = `Simulated Data - Player`,
+    `Sim MVP` = `Simulated Data - MVP Index`,
+    `Hist Player` = `Historical Data - Player`,
+    `Hist MVP` = `Historical Data - MVP Index`
+  )
+
 tiny_combined_table <- tt(combined_table)
 
-print(combined_table)
+```
+
+```{r}
+#| echo: false
+
 print(tiny_combined_table)
-
 ```
@@
 
-print(mvp_table)
+mvp_table <- mvp_table %>% select(-Source)
+
+tiny_mvp_table <- tt(mvp_table)
+
+print(tiny_mvp_table)
 ```
@@
 
+Real_MVP_Data_f <- Real_MVP_Datas %>% select(-FG., -X3P., -FT., -WS.48, -Voting, -Lg, -Tm, -G, -MP)
+
 write.csv(Real_MVP_Datas, "../data/02-analysis_data/Real_MVP_Datas.csv", row.names = FALSE)
 
-print(Real_MVP_Datas)
+tiny_real_mvp <- tt(Real_MVP_Data_f)
+print(tiny_real_mvp)
 ```
@@
 #| echo: false
+#| 
 lm_summary <- summary(lm_model)

### 2024-12-14 paper/paper.qmd
Formatting Edits

@@
 #| label: fig-mvpindex
-#| fig-cap: "MVP index of of Historical and simulated MVPs"
+#| fig-cap: "MVP index of Historical and simulated MVPs"
 
@@
   labs(
-    title = "Scatter: Year vs. MVP Index",
     x = "Year",
@@
 #| label: fig-pts
-#| fig-cap: "MVP PPG over Time"
+#| fig-cap: "PPG of Historical and simulated MVPs"
 
@@
 #| label: fig-reb
-#| fig-cap: "MVP RPG over Time"
+#| fig-cap: "RPG of Historical and simulated MVPs"
 
@@
 #| label: fig-ast
-#| fig-cap: "MVP ASG over Time"
+#| fig-cap: "APG of Historical and simulated MVPs"
 
@@
 This suggests that MVPs in the modern NBA are increasingly likely to be selected if they are well-rounded players, athletes who not only score points and rebound effectively but also contribute as unselfish teammates by creating scoring opportunities through their passing.
+
+\newpage
 
@@
 
-## Weaknesses and next steps {#sec-weak}
+## Weaknesses and Next Steps {#sec-weak}
 
@@
 These simulations highlight the significant role that methodological choices play in shaping MVP selections, showing that even small adjustments can result in different winners.
-By simulating different voting scenarios, it becomes clear that the current system is sensitive to how votes are counted and weighted, underscoring the need for transparency and careful consideration of methodology.
+By simulating different voting scenarios, it becomes clear that the current system is sensitive to how votes are counted and weighted, emphasising the need for transparency and careful consideration of methodology.
 

### 2024-12-14 paper/paper.qmd
Small edits to spelling and grammar

@@
 However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas.
-As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate, as well as an $R^{2}$ of $R^{2} = 0.8992$, with a corresponding $R = 0.9483$.
+As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate, as well as an $R^{2}$ score of $R^{2} = 0.8992$, with a corresponding $R = 0.9483$.
 
@@
 Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents.
-However, their inconsistency can leave teams vulnerable during less-publicized matchups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."
+However, their inconsistency can leave teams vulnerable during less-publicized match ups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."
 
@@
 Next, team records are also incredibly important for MVP, if a player played incredibly one season but barely made the playoffs, or even did not at all, the player likely should not be receiving the MVP award, as how much worse could said team get if they weren't that good with the so called "best" player in the league.
-Unfortunately we were notable to inclue team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed.
+Unfortunately we were notable to include team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed.
 Unfortunately this data was not included in either type of data set that we sourced from [@bbref], and due to time constraints downloading potentially dozens if not hundreds additional data sets did not seem to be the best use of time in our case.
@@
 
-Our next steps, would like be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurate predict the model.
-We would also like to revisit this paper tat the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.
+Ideally, our next steps, would be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurately predict the model.
+We would also like to revisit this research at the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.
 
